---
title: "A framework for a kinference vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"A framework for a kinference vignette"}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Kinference vignette

## First, the walkthrough of 'kinference' provided to SB by Paige

```{r}
library( kinference) ## don't overwrite it with the one from the Rmvb repo!`
```

Install/update packages from Mark's repository

`# install.packages(repos='https://markbravington.github.io/Rmvb-repo')`

`# update.packages(repos='https://markbravington.github.io/Rmvb-repo')`

```{r}
    library( mvbutils)
    library( atease)
    library( xtable)
    library( mixtools)
    library( Oarray)
    library( Oarray.support)
    library( vecless)
    library( tools)  # need or else get error "could not find function "md5sum""
    library( gbasics)  #has sqr, logit and inv.logit, ridder
    library( genocalldart)
    ## library(handy2) ## not requred under the current build of kinference
```

## To see what functions are in a package:

`ls("package:kinference")`

Load object "geno2017" that contains the genotype calls for all of the samples processed
by DArT prior to 2018

```{r}
load('~/Data/kinference/geno2017.rda')
geno_opt <- "2017"
```

Note that geno2017 was created by several calls to "load_whopper". The R code is as follows:

```{r}
#g1<- load_whopper('Report-DTu17-2599_part1.csv', which_rows = c(
#         well = 1,
#         dartjob = 2,
#         ourplate = 3,
#         fishname = 4,
#         fishtot = 5),
#     gobj=g0[0,],
#     cutoff_npoly=0.5, dropped_fish_file='DTu17-2599_part1_dropped_fish.txt',max_fish=2500)
#g2<- load_whopper('Report-DTu17-2599_part2.csv', which_rows = c(
#         well = 1,
#         dartjob = 2,
#         ourplate = 3,
#         fishname = 4,
#         fishtot = 5),
#     gobj=g1,
#     cutoff_npoly=0.5, dropped_fish_file='DTu17-2599_part2_dropped_fish.txt',max_fish=2500)
#g3 <- load_whopper('Report-DTu17-2711.csv', which_rows = c(
#         well = 1,
#         dartjob = 2,
#         ourplate = 3,
#         fishname = 4,
#         fishtot = 5),
#     gobj=g2,
#     cutoff_npoly=0.5, dropped_fish_file='DTu17-2711_dropped_fish.txt',max_fish=2500)
#geno2017<- g3
```

Each call:
 - reads in the data from a DArT count csv file (as specified)
 - determines how many polyploid loci each fish has, and gets rid those with too many(set by an
   argument to the load_whopper function)
 - determines the genotypes based on the counts for every fish/loci to the 6-way level (e.g. AA/BB/AB/AO/BO/OO);
  it does so based on criteria/thresholds that were determined previously for each of the loci, and which
are stored in the attributes of an object called "g0"
 - the first call to load_whopper needs passed the object "g0", then each consecutive call gets passed the
object created from the previous call and appends the new genotype data to it.


## Run fish QC checks

Save all figures in a pdf:

`pdf('fish_QC_plots.pdf')`

(i) Run original hetz check (that automatically dumps fish)

```{r}
multhresh <- 0.001

geno_okhetz <- dump_badhetz_fish(geno2017, multhresh_badhetz_fish = nrow(geno2017)*multhresh,badhetz_hist_pars=list(nclass=150,xlim=c(0,600)), 'hetzminoo') 
dim(geno_okhetz)
```

(ii) Get rid of duplicates (Mark does this first, but if you don't get rid of the really bad hetz fish, then it's not as clear/clean which fish are duplicates):

```{r}
dups <- find_duplicates(geno_okhetz,max_diff_genos=150)
rev(sort(dups$ndiff))[1:20]
droppies <- drop_dups_pairwise_equiv(cbind(dups$i,dups$j))
# Mark doesn't get rid of the 3 remaining controls, but I think we probably should
controls <- unique(c(grep('CR',geno_okhetz@info$Our_sample,ignore.case=T),grep('control',geno_okhetz@info$Our_sample,ignore.case=T)))
controls[controls%not.in%droppies]
droppies <- unique(c(droppies,controls))
geno_nodups <- geno_okhetz[ -droppies, ]
dim(geno_nodups)
```


(iii) Run the expected likelihood test and get rid of

```{r}
ilglk <- ilglk_geno(geno_nodups)
ilglk_lb <- (-1550)  ## SB: arbitrary lower bound here?
abline(v=ilglk_lb,col=6,lwd=2)
geno_nodups_goodfish <- geno_nodups[ilglk>ilglk_lb,]
dim(geno_nodups_goodfish)
```

(iv) Run the "fancier" hetzminoo check (no automatic dumping)

```{r}
hetz.rich<- hetzminoo_fancy(geno_nodups_goodfish, 'rich')
hetz.poor<- hetzminoo_fancy(geno_nodups_goodfish, 'poor',hist_pars=list(nclass=100))
poor_lb <- 0.196
abline(v=poor_lb,col=6,lwd=2)
geno_nodups_gooder <- geno_nodups_goodfish[hetz.poor>poor_lb,]
# Note that "gooder" was Mark's choice of terminology!
dim(geno_nodups_gooder)
```

`dev.off()`


Check that the loci are still looking okay for these fish

```{r, eval = FALSE}
pdf(paste('check6and4.pdf'))
check6and4(geno_nodups_gooder,c(1e-4,1e-3))  ## not shown
dev.off()
```


## FIND POPS

for finding POPs, the "good fish" version is sufficient ... don't need to go use the "gooder" version which does more drastic culling of fish based on hetzmino_fancy (perhaps required for finding HSPs)

```{r, eval = FALSE}
adults<- grep('SbIN',geno_nodups_goodfish@info$Our_sample,ignore.case=T)
juvs <- grep('SbPL',geno_nodups_goodfish@info$Our_sample,ignore.case=T)

pops <- find_POPs(geno_nodups_goodfish, subset1=adults, subset2=juvs, one_in_X_eta=1e6,
                  keep_n=1000,
                  nbins=50) ## find_POPs has a non-zero runtime, so is not evaluated in the vignette
## save(pops,file=paste('pops2018_',geno_opt,'.rda',sep=''))

pdf(paste('pops_histo_',geno_opt,'.pdf',sep=''))
hist(pops$wpsex,nc=40)
abline(v=0.045,col=2,lwd=2) 
dev.off()
```


### How many look like POPs?

```{r, eval = FALSE}
sum(pops$wpsex<0.045)
```


## FIND HSPS

For finding HSPs,
(a) start with the "good fish" version is sufficient, then
(b) compare with the "gooder" version which does more drastic culling of fish based on hetzmino_fancy

```{r}
fish_opt <- "good"
#fish_opt <- "gooder"
if(fish_opt=="good") geno_for_hsps<- geno_nodups_goodfish
if(fish_opt=="gooder") geno_for_hsps<- geno_nodups_gooder
```

reduce the dataset to only juveniles

```{r, eval = FALSE}
is.juv <- grepl('SbPL',geno_for_hsps@info$Our_sample,ignore.case=T)
geno_juves<- geno_for_hsps[is.juv,]
dim(geno_juves)

fname=paste('hsps2018_',geno_opt,'_',fish_opt,'.rda',sep='')
if(file.exists(fname)) load(fname) ## if already generated, don't repeat next slow step.
if(!file.exists(fname)) {
  # run find_HSPs (need to run prepare_PLOD_SPA first, or else get error that Kenv is missing)
  geno_juves <- prepare_PLOD_SPA(geno_juves)
    hsps <- find_HSPs(geno_juves, one_in_X_eta=1e6, keep_n= 1000,
                      bins=seq(-150,250,5)) ## find_HSPs is also not run in vignette owing to runtime
##    hsps_dep <- find_HSPs_deprecated(geno_juves, one_in_X_eta=1e6, rough_n_pairs_to_keep= 200,
##                                  bins=seq(-150,250,5))

##  save(hsps,file=fname)
}

hist(hsps$PLOD[hsps$PLOD > 0], nc = 40)
```



## The original contents of 'vignette':

    library(gbasics)
    #library(kinference)
    library(devtools)
    load_all("~/current/kinference")

required to inspect @ elements

`library(atease)`


need to impress on people here that putting the RIGHT data.frame in the right function is essential or BAD THINGS will happen, also keeping track of what stage a given data.frame is at

    `load("dave-practice.rda")`

    `load("fsps_for_dave.rda")`

    `dave <- rbind( dave, more_for_dave)`

## note on adding in extra data and the subset arguments

what are we **doing** here? Removing totally unlikely fish?
with other data sets need to fiddle with the xlims?

`ill <- ilglk_geno(dave)`

## need a hetzminoo_fancy thing here? Need to setup prepare_PLOD_SPA too?

get duplicates
from manual max_diff_genos "Try increasing this from say 10 upwards"
nice little story here where if you set to 40 you get a spurious result
x axis is "number of loci different" -- say something about QC

`dupes <- find_duplicates(dave, max_diff_genos=100)`

`hist(dupes$ndiff)`

get the indices to drop
CHECK THIS function

`droppies <- drop_dups_pairwise_equiv(dupes[,2:3])`

`dave_nodupes <- dave[-droppies,]`

no duplicates any more

`nodupes <- find_duplicates(dave_nodupes, max_diff_genos=100)`

want to look at "non-technical duplicates" here too?

`drop_dups_pairwise_equiv(..., want_groups=TRUE)`

look at `$Our_sample`

## find POPs

`pops <- find_POPs(dave_nodupes, one_in_X_eta=1e3, nbins=50, keep_n = 50000)`

`hist(pops$bigs$wpsex)`

## something about nABOO here?

`#all_pops <- 0.051`

## find HSPs

`dave_nodupes <- prepare_PLOD_SPA(dave_nodupes)`

`is_juve <- toupper(substring(dave_nodupes@info$Our_sample, 3, 4))=="PL"`

`dave_juves <- dave_nodupes[is_juve, ]`

## bins found using predict_hsp_util???

`dd <- find_HSPs(dave_juves, bins = seq(-120, 120, by = 5), keep_thresh = -5, eta = 10, keep_n=1000)`

usually bump up number of classes

    hist(dd$PLOD, nc=30)
    #abline(dd$mean thing)

could remove i and j s for POP and FSP pairs

## histogram

`plot(dd@bins, dd@n_PLODs_in_bin)`

## log-histogram

`plot(dd@bins, log(dd@n_PLODs_in_bin), type="b")`

HTPs around 0 in this?

for get chain

`table(dd[dd$PLOD >0 & dd$PLOD <30,][,c("i","j")])`

## getting FSPs

HSP pairs in a matrix 2 cols

`HSPs <- as.matrix(dd[, 2:3])`

`pp <- find_FSPs_from_HSPs(dave_juves, HSPs)`

## remove FSPs from HSPs

calculate the "upper variance" not including the FSPs

`variance sum((PLOD - EPLOD_HSP)^2)`

HTP/HCP
UP -- find_HSPs set one_in..
HCP/HTP -- 

## false negative rate

## get_chain stuff here?

## example table of numbers of HSPs per year
## number of comparisons


