---
title: "Inferring kin relationships with kinference"
author: S M Baylis
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{inferring kin relationships with kinference}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Package `kinference` allows users to use SNP genotype data from
biological samples to infer kin relationships between pairs of
samples. In this vignette, we demonstrate how to use the core
functions of package `kinference` to perform kin-finding and the
necessary data-cleaning steps that lead up to kin-finding (section 1),
and show a handful of canned examples of the sorts of things that can
go wrong (section 2).

# Section 1: data-cleaning and kin-finding

Here, we step through kin-finding with a well-behaved, well-cleaned
dataset in `snpgeno` format. We will use the `dropbears` dataset,
included with package `kinference`. We'll show the standard checks
that should _always_ be done, but in this case the dataset is already clean
and won't require any changes. To see a dataset in need of cleaning,
see *section 2*.

A dataset for kinference will need all of its samples drawn from a
single, well-mixed population, with no duplicate samples, samples from
another population (or species!), or cross-contaminated or degraded
samples. It also needs all loci to represent well-behaved single loci
in Hardy-Weinberg proportions, with no allele dropout, gene
duplications, loci under (strong) selection, or mistakenly-identified
loci from hard-to-align regions such as microsatellites.

## checking loci for HWE proportions

First, we check whether our loci look like they meet our expectations
by testing each locus for HWE proportions across all samples, using
function `check6and4`.

```{r dropbear6and4}
## load the data
data(dropbears)
library(kinference)
## check loci for HWE proportions
pvals <- check6and4(dropbears, thresh_pchisq_6and4 = c(0.01, 0.0001))
```

`check6and4` calculates Chi-square P-values for each the observed
genotypes at each locus across all samples under a null hypothesis of
Hardy-Weinberg proportions, and plots the observed and expected counts
of each genotype for each locus. Lower P-values are indicate
worse-looking loci. Our P-values are mostly high, and there is a tight
agreement between the expected and observed counts for every
genocall. These loci are fine.

## Checking for duplicate samples

Next, we should make sure we don't have any duplicate
samples. Duplicate samples can be intentional, but surprisingly many
datasets contain accidental duplicates. Even if you think your lab
skills are immaculate, it's worth checking.

```{r dropbearFindDups}
## look for duplicates in 'dropbears' and keep any pairs with
## less than 50 different genotypes:
dups <- find_duplicates(dropbears, max_diff_loci = 50)
```

`find_duplicates` compares all pairs of samples, and records any pairs
with fewer than a specified number of differences. You should specify
that number yourself, based on the number of loci in your dataset and
your expected genotype mis-call rate. `find_duplicates` also generates
a plot showing the density of pairs with _n_ different genotypes. In
this case, it doesn't look like we have any duplicates.

## Checking for samples from another population or species

Next, we should make sure we don't have any 'tourists' from another
population (or species!) in our dataset, using function `ilglk_geno`.

```{r dropbearIlglk}
ilglk <- ilglk_geno(dropbears)
```

For each sample, `ilglk_geno` calculates the likelihood that its
observed genotypes could be drawn at random from a population with
allele frequencies estimated from all of the samples. It also plots an
ideal density distribution for that likelihood statistic, if all
allele frequencies were perfectly estimated from the data. Samples
that are outlying to the left of the rest of the samples are probably
tourists from another population, or maybe mis-identified to species,
and should be removed. You may notice that the idealised density
distribution doesn't _quite_ match our observed density - the observed
density is slightly right-shifted compared to the idealised distribution. In
practice this is pretty common, because of linkage and imperfect
estimation of allele frequencies from the data. This is expected, and
a small amount of mismatch (like here) is fine.

## Checking for cross-contaminated or degraded samples

Next, we should check for cross-contaminated samples, using function
`hetzminoo_fancy`. Cross-contaminated samples tend to have high
apparent heterozygosity, and so they look similar to each other, and
can be wrongly interpreted as kin. We also need to run a boring
book-keeping calculation using the function `hsp_power`.

```{r dropbearHetzRich}
## a boring book-keeping calculation that needs to happen before checking for contamination:
dropbears <- hsp_power(dropbears, k = 0.5)
## check for cross-contaminated samples
hetzrich <- hetzminoo_fancy(dropbears, 'rich')
## check for degraded samples
hetzrich <- hetzminoo_fancy(dropbears, 'poor')
```

For each sample, `hetzminoo_fancy` looks at the distribution of
heterozygotes and each type of homozygote across all loci, compared to
the expected distributions of heterozygotes and homozygotes. It can
compute two slightly-different statistics: with `taget = 'rich'` it is
optimised to find cross-contaminated samples, and with `target =
'poor'` it is optimised to find degraded samples. It also calculates
an expected distribution for each statistic. In either case, outlying
samples should be removed - but these samples all look fine.

## Finding and classifying kin

Finally, we get to calculate our kin-finding statistics and classify
kin. Our workflow goes something like:

  * find 'strong' kin, ideally first-order kin (i.e., parent-offspring
    or full-siblings), but excluding all unrelated pairs but
    potentially including some weaker kin such as half-siblings.
  * within the 'strong' kin, differentiate parent-offspring pairs from
  full-sibling pairs (and any weaker kin), and record them so they can
  be stripped from subsequent analysis.
  * find weaker kin, i.e., second-order (half-sibling pairs,
  grandparent-grandoffspring pairs, or full aunt-niece pairs) and
  third-order (half aunt-niece, full cousins), and possibly some
  fourth-order kin
  * estimate a mixture distribution for second-order and third-order
  kin, and use it to find a set of second-order kin and estimate the
  false-negative rate for that set.

Let's start by identifying our first-order kin

## Find Parent-offspring pairs

```{r dropbears_findPOPs}
## set the kin palette
palette(c("#0D0887FF", "#48039FFF", "#7401A8FF", "#9D189DFF", "#BF3984FF",
          "#DA596AFF", "#EE7B51FF", "#FBA238FF", "#FCCE25FF"))

## find strong kin using find_POPs
strongKin <- find_POPs(dropbears, keep_thresh = 0.6)
hist(strongKin$wpsex, breaks = 50)
lines(strongKin@bins,diff(c(0,pnorm(strongKin@bins, mean = strongKin@mean_UP,
                    sd = sqrt(strongKin@var_UP)) * choose(nrow(dropbears), 2))),
      lwd = 2, col = 5) ## col = 5 for UPs
```

The `wpsex`statistic calculated by `find_POPs` has an expectation of zero for
true parent/offspring pairs with no genotyping error, and we see a
spike of about 20 pairs near zero, and a few more bumps of kin. We
have a theoretical mean and variance of this statistic for unrelated
pairs, which we've plotted in purple. Based on that distribution, we
would expect no unrelated pairs with a `wpsex` score less than 0.055,
so let's keep all pairs with `wpsex < 0.055` as our set of 'strong'
kin, and start separating them out.

```{r dropbears_split_FSPs_POPs}
UPcut <- 0.055  ## keep only the clear kin from strongKin
## and calculate a statistic designed to separate full-siblings from parent-offspring pairs
## par(mfrow = c(3,1))

## note that we have to specify a genotyping error rate here - but the
## calculation is quite robust to mis-specification. 0.01 is probably
## a safe default if you don't have a good estimate for your dataset -
## and if your true error rate is much higher than that, you probably
## won't be able to use your dataset for kin-finding anyway.
splitPOPs_1 <- split_FSPs_from_POPs(dropbears, strongKin[strongKin$wpsex < UPcut,], gerr = 0.01)
hist(splitPOPs_1$PLOD_FP, breaks = 50, main = "", xlab = "PLOD_FP")

POPs <- splitPOPs_1[splitPOPs_1$PLOD_FP < 0,] ## our first set of definite kin pairs
POPs$ij <- paste(POPs$i, POPs$j) ## a pair ID
```

We now have a set of known Parent-offspring pairs. The remaining
first-order kin must be full-sibling pairs, so let's find them

## Find Full-sibling pairs

```{r findHSPs}
## find HSPs-or-stronger using find_HSPs
HSPs_1 <- find_HSPs(dropbears, keep_thresh = -10)
PLOD_loghisto(HSPs_1)
## note the distribution of the PLOD - there is a large bump of
## unrelated pairs, with mean and variance exactly where theory
## predicts them to be, and bumps for second-order (half-sibling) and
## first-order (full sibling and parent-offspring) pairs centred
## exactly where theory expects. This is a good dataset.
```

Any first-order kin that _aren't_ POPs must be FSPs. We can see
there's a clear gap between the half-sibling pair bump and the
first-order kin bump, so we can take the first-order kin from this
statistic, subset out the parent-offspring pairs, and be left with the
full-siblings.

```{r assignFSPs}
POPsOrFSPs <- HSPs_1[HSPs_1$PLOD > 110,]
POPsOrFSPs$ij <- paste(POPsOrFSPs$i, POPsOrFSPs$j) ## a pair ID to match the one in POPs
FSPs <- POPsOrFSPs[!POPsOrFSPs$ij %in% POPs$ij,] ## find our full-sibling pairs
```

Now, we have our set of parent-offspring pairs, our set of
full-sibling pairs, and a set of pairs tentatively labelled as
'half-siblings'. But that 'half-sibling' set may (in fact, will!)
contain weaker kin such as thiatic pairs ('thiatic' is the
gender-neutral term that includes aunt-niece, uncle-niece,
aunt-nephew, and uncle-nephew pairs) or cousin pairs. We need to trim
out the first-order kin from this set, then decide on a cutoff between
half-siblings and thiatics that allows us to keep a large proportion
of our sibling pairs, with a very low, known, rate of false positives.

```{r splitHSPsFromThiatics}
HSPs_1$ij <- paste(HSPs_1$i, HSPs_1$j)
HSPs_2 <- HSPs_1[!HSPs_1$ij %in% c(POPs$ij, FSPs$ij),] 
## plot the empirical PLOD scores with the expected distribution
## of the PLOD for half-sibling pairs
HSP_histo(HSPs_2, lb = 0, fullsib_cut = 110, bin = 3)

autopick_threshold(dropbears, kin = HSPs_2, fitrange_PLOD = c(0, 110),
                   FPtol_pairs = 1, use4th = TRUE, selection = "ML")
```




```{r probablyNotKeepingThis}
## ## HSP cutoff with few false negatives, some false positives
 ## HSPcut <- 10
 ## abline(v = HSPcut, col = "red", lwd = 2)
 ## HSPs_1 <- HSPs_1[HSPs_1$PLOD > HSPcut,]
 ## HSPs_1$ij <- paste(HSPs_1$i, HSPs_1$j)
 ## ## exclude the FSPs
 ## HSPs_1 <- HSPs_1[!HSPs_1$ij %in% FSPs_1$ij,]
 ## ## exclude the (empty) POPs
 ## HSPs_1 <- HSPs_1[!HSPs_1$ij %in% POPs_1$ij,]
 ## ### scrub the thiatics out of the HSPs
 ## ### first check: do we have an HSP bump, centred where it ought to be?
 ## ### _only_ do this for HSPs. No FSPs or POPs!
 ## HSP_histo(HSPs_1, lb = 10, fullsib_cut = 130)
 ## ### yes; we have an HSP bump, and its mean is where it ought to be
 ## PLODs_over_mean <- HSPs_1[HSPs_1$PLOD > HSPs_1@mean_HSP,]
 ## var_HSP <- mean((PLODs_over_mean$PLOD - HSPs_1@mean_HSP)^2)
 ## ## _eyeball_ a cutoff such that no more than one HTP should exceed it
 ## myEye <- 55
 ## ## what proportion are going to be below myEye if the mean and var are
 ## ## what we calculated?
 ## fneg_eye <- pnorm(myEye, mean = HSPs_1@mean_HSP, sd = sqrt(var_HSP))
 ## ## Or, if we want to be defensible and fancy about it...
 ## ## There is an 'objective' method.
 ## var_HTP <- var_PLOD_kin(SHS_g@locinfo, emp_V_HSP = var_HSP, kin_true = "HTP")
 ## ### the conservative approach is to pick the bigger of the possible
 ## ### variances from var_HTP for HTPs. We think there's a bug in Vx; but
 ## ## in principle, take the bigger one.
 ## bigVar <- max(var_HTP[cq(V0.HTP, Vx)])
 ## ## now, guess-timate how many HTPs there might be in the whole
 ## ## dataset.
 ## ### very conservative is 3x the number of HSPs (because the long-run
 ## ## truth is 2x). So use 6x the number of rows in PLODs_over_mean
 ## rough_n_HTPs <- 6*nrow(PLODs_over_mean)
 ## ## find a PLOD value above which there oughtn't to be more than one
 ## ## HTP.
 ## myBrain <- qnorm((1/rough_n_HTPs), mean = 0, sd = sqrt(bigVar), lower.tail = FALSE)
 ## ## given that cutoff, what's our false-negative rate for HSPs?
 ## fneg_brain <- pnorm(myBrain, mean = HSPs_1@mean_HSP, sd = sqrt(var_HSP))
 ## ### finally, trim HSPs to only those above your chosen cutoff
 ## HSPs_1 <- HSPs_1[HSPs_1$PLOD > myBrain,]
```



# Section 2: examples of things that can go wrong

Go through early school shark for this?


```{r}
## data(dropbears)
## pvals <- check6and4(dropbears, thresh_pchisq_6and4 = c(0.01, 0.0001))
```

The argument thresh_pchisq_6and4 sets two cutoffs
for 'bad-ish' and 'really bad' loci. In this example, we have loci
with ... . We have nnnn loci, so this is more than we would expect:
we've got some bad loci here, and should remove them
