---
title: "Inferring kin relationships with kinference"
author: S M Baylis
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inferring kin relationships with kinference}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

Package `kinference` allows users to use SNP genotype data from
biological samples to infer kin relationships between pairs of
samples. In this vignette, we demonstrate how to use the core
functions of package `kinference` to perform kin-finding with a
perfect dataset (section 1), cleaning a 'typical' dataset to get it to
work for kinfinding (section 2), and some examples showing the warning
signs of Things Going Wrong (section 3).

# Section 1: kin-finding

Here, we step through kin-finding with a well-behaved, well-cleaned
dataset in `snpgeno` format. We will use the `dropbears` dataset
included with package `kinference`. Some standard checks should
_always_ be done, but the `dropbears` dataset is already clean and
won't require any changes. To see a dataset in need of cleaning (like
yours!), see **Section 2**.

A dataset for kinference will need all of its samples drawn from a
single, well-mixed population, with no duplicate samples, samples from
another population (or species!), or cross-contaminated or degraded
samples. It also needs all loci to represent well-behaved single loci
in Hardy-Weinberg proportions, with no allele dropout, gene
duplications, loci under (strong) selection, or mistakenly-identified
loci from hard-to-align regions such as microsatellites.

## Plot an overview of the kin-space

```{r dropbears_overview}
library(kinference)
library(atease)
data(dropbears)
kincol <- kinPalette() ## set the kin palette and store colour names for later

## a couple of boring but necessary data-prep steps:
dropbears <- kin_power(dropbears, k = 0.5)
dropbears <- prepare_PLOD_SPA(dropbears)

plod_HU <- find_HSPs(dropbears, keep_thresh = -10)
PLOD_loghisto(plod_HU)
```

This plot shows observed log-densities of a kin-finding statistic as
the black stair-step lines (the statistic is optimised for splitting
half-siblings from unrelated pairs). It also shows some vertical lines
at the the expected means for various classes of kin, and the expected
distribution of unrelated pairs as a smooth purple curve. This is an
excellent-looking kin space - we'll definitely be able to find kin
here.

## Finding and classifying kin

From here, there are many potential workflows for finding kin. The
workflows can depend on demographic information (e.g., one might look
for sibling pairs only amongst samples of young animals caught at a
similar time, to reduce confusion from aunt-niece pairs and
parent-offspring pairs; and look for parent-offspring pairs only among
pairs with a sizeable birth-year gap). Here, we show one potential
workflow for kin-finding that does not use demographic information. It
goes:

  * find 'strong' kin, ideally first-order kin (i.e., parent-offspring
    or full-siblings), but excluding all unrelated pairs but
    potentially including some weaker kin such as half-siblings
  * within the 'strong' kin, differentiate parent-offspring pairs from
  full-sibling pairs (and any weaker kin), and record them so they can
  be stripped from subsequent analysis.
  * find weaker kin, i.e., second-order (half-sibling pairs,
  grandparent-grandoffspring pairs, or full aunt-niece pairs) and
  third-order (half aunt-niece, full cousins), and possibly some
  fourth-order kin
  * estimate a mixture distribution for second-order and third-order
  kin, and use it to find a set of second-order kin and estimate the
  false-negative rate for that set.

Let's begin.

## Find 'strong' kin and split out parent-offspring pairs

```{r dropbears_findPOPs}
## find strong kin using find_POPs
strongKin <- find_POPs(dropbears, keep_thresh = 0.6)
## `keep_thresh` of 0.6 is arbitrary. The theoretical lower limit is
## 0, and in principle keep_thresh should be set at a value where
## unrelated pairs comprise substantial proportion of the density, as
## shown in this graph:
hist(strongKin$wpsex, breaks = 50)
lines(strongKin@bins,diff(c(0,pnorm(strongKin@bins, mean = strongKin@mean_UP,
                    sd = sqrt(strongKin@var_UP)) * choose(nrow(dropbears), 2))),
      lwd = 2, col = kincol['UP']) 
## The curved line gives predicted distribution for Unrelated Pairs.
## There is also a likelihood-based statistic for POP-finding, but it's
## new and a little fiddly. You would run it with this code:
## gerr next is a _rough_ idea of genotyping error rate; results should
## not be highly sensitive to it
## strongKin2 <- find_POPs_lglk(dropbears, gerr = 0.03, keep_thresh = -5)
## HSP_histo(strongKin2, lb = -5, fullsib_cut = 0)
## PLOD_loghisto(strongKin2)
```

The `wpsex`statistic calculated by `find_POPs` has an expectation of zero for
true parent/offspring pairs with no genotyping error, and we see a
spike of about 20 pairs near zero, and a few more bumps of kin. We
have a theoretical mean and variance of this statistic for unrelated
pairs, which we've plotted in purple. Based on that distribution, we
would expect no unrelated pairs with a `wpsex` score less than 0.055,
so let's keep all pairs with `wpsex < 0.055` as our set of 'strong'
kin, and start separating them out.

```{r dropbears_split_FSPs_POPs}
UPcut <- 0.055  ## keep only the clear kin from strongKin
## and calculate a statistic designed to separate full-siblings from parent-offspring pairs
## par(mfrow = c(3,1))

## note that we have to specify a genotyping error rate here - but the
## calculation is quite robust to mis-specification. 0.01 is probably
## a safe default if you don't have a good estimate for your dataset -
## and if your true error rate is much higher than that, you probably
## won't be able to use your dataset for kin-finding anyway.
splitPOPs <- split_FSPs_from_POPs(dropbears, strongKin[strongKin$wpsex < UPcut,], gerr = 0.01)
hist(splitPOPs$PLOD_FP, breaks = 50, main = "", xlab = "PLOD_FP")
## Colours: kincol was set by call to kinPalette()
abline(v = splitPOPs@E_FSP, col = kincol['FSP'], lwd = 2) 
abline(v = splitPOPs@E_POP, col = kincol['POP'], lwd = 2)
legend("topright", legend = c("POP expectation", "FSP expectation"), lty = c(1,1), lwd = c(2,2),
       col = kincol[ c('POP', 'FSP')], bg = "white")
## the clump between -50 and -100 is POPs; the rest are a mix of FSPs, HSPs, and possibly weaker kin.

POPs <- splitPOPs[splitPOPs$PLOD_FP < 0,] ## our first set of definite kin pairs

POPs$ij <- paste(POPs$i, POPs$j) ## a pair ID
```

We now have a set of known Parent-offspring pairs. The remaining
first-order kin must be full-sibling pairs, so let's find them.

## Find Full-sibling pairs

```{r findHSPs}
## find HSPs-or-stronger using find_HSPs
HSPs_1 <- find_HSPs(dropbears, keep_thresh = -10)
PLOD_loghisto(HSPs_1)
## note the distribution of the PLOD - there is a large bump of
## unrelated pairs, with mean and variance exactly where theory
## predicts them to be, and bumps for second-order (half-sibling) and
## first-order (full sibling and parent-offspring) pairs centred
## exactly where theory expects. This is a good dataset.
```

Any first-order kin that _aren't_ POPs must be FSPs. We can see
there's a clear gap between the half-sibling pair bump and the
first-order kin bump, so we can take the first-order kin from this
statistic, subset out the parent-offspring pairs, and be left with the
full-siblings.

```{r assignFSPs}
POPsOrFSPs <- HSPs_1[HSPs_1$PLOD > 110,]
POPsOrFSPs$ij <- paste(POPsOrFSPs$i, POPsOrFSPs$j) ## a pair ID to match the one in POPs
FSPs <- POPsOrFSPs[!POPsOrFSPs$ij %in% POPs$ij,] ## find our full-sibling pairs
```

Now, we have our set of parent-offspring pairs, our set of
full-sibling pairs, and a set of pairs tentatively labelled as second-order kin. "Second-order kin" comprise half-sibling, full-thiatic, and grandparent-grandchild pairs ('thiatic' is a succinct bidirectional gender-neutral term that comprises aunt-niece, uncle-niece, aunt-nephew, and uncle-nephew kinships): we cannot split those genetically, but in most CKMR applications, half-siblings will be commonest on demographic grounds (details out of scope here). However, the bigger issue is that the set of tentative second-order kin may well be contaminated by more-distant kin, mainly third-order (half-thiatic and great-grandparent-grandchild pairs) and fourth-order (e.g. full first-cousin pairs). We need to trim
out the 1st-order kin from this set, then decide on a cutoff between
2nd- and 3rd-orders that allows us to keep a large proportion
of our half-sibling pairs, with a known and very low rate of false positives.

To achieve this, we use a mixture decomposition of the half-sibling,
thiatic, and cousin pairs, set to choose a PLOD threshold that only a
specified number of half-thiatic or cousin pairs will exceed.

```{r splitHSPsFromThiatics}
HSPs_1$ij <- paste(HSPs_1$i, HSPs_1$j)
HSPs_2 <- HSPs_1[!HSPs_1$ij %in% c(POPs$ij, FSPs$ij),] 
## plot the empirical PLOD scores with the expected distribution
## of the PLOD for half-sibling pairs
thresh <- autopick_threshold(dropbears, kin = HSPs_2, fitrange_PLOD = c(0, 110),
                   FPtol_pairs = 1, use4th = TRUE, selecto = "ML", plot_bins = 5)
```

The plot from `autopick_threshold` shows our observed distribution of
the PLOD_HU (a statistic optimised for splitting half-siblings from
unrelated pairs) alongside the component curves from a mixture
decomposition model aiming to split the PLOD_HU into second-order
(e.g., half-siblings), third-order (e.g., half-thiatics), and
fourth-order (e.g., half-cousins) pairs and the combined density of
those component curves. The observed PLOD_HU distribution should
closely match the black curve. It has then picked a cutoff to split
HSPs from all weaker kin-types, with only a small number of
false-positive half-siblings. In this example, we've set that number
of false-positives to 1, with the argument `FPtol_pairs = 1`.

With that cutoff, we can define our set of HSPs and the estimated
false-positive and false-negative rates of that set:

```{r assignHSPs}
HSPs <- HSPs_2[HSPs_2$PLOD > thresh,]
fpos_rate <- thresh@FPtol_pairs / nrow(HSPs) ## approx 2% of our 'HSPs' are false-positives
fneg_rate <- thresh@info["Pr_FNeg"] ## we're missing about 0.5% of HSPs to false-negatives
nrow(HSPs) ## 52 inferred half-sibling pairs
```

## Link kin pairs to their sample covariate data

Now, we have three sets of kin pairs, but they're not linked to sample
covariate data like an individual ID - individuals are identified by
their row numbers in the dataset used for analysis, but for any
subsequent analysis, some additional sample information will almost
certainly be useful. Let's associate the pairs with their sample
covariates.

```{r HSPmetadata}
HSPs <- get_pair_covars(dropbears, HSPs)
POPs <- get_pair_covars(dropbears, POPs)
FSPs <- get_pair_covars(dropbears, FSPs)
splitPOPs <- get_pair_covars(dropbears, splitPOPs)

## covariate data are stored in @i_covars and @j_covars:
HSPs@i_covars[1:5,]
HSPs@j_covars[1:5,]
```

# Section 2: cleaning the bluefin dataset

Your data will almost certainly need cleaning before you can do the
kin-finding in Section One. To show how (and, in section 3, _why_),
we'll use the `bluefin` dataset included in package `kinference`.

## checking for badly-behaved loci

The `bluefin` dataset has 1510 loci, and in an abstract perfect
dataset, we would expect their P-values to be uniformly distributed
over the interval [0,1] - so we would _expect_ about 0.151 of them to
have a P-value less than 1e-4. 

```{r bluefin6and4}
data(bluefin)
pvals <- check6and4(bluefin, thresh_pchisq_6and4 = c(1e-4, 1e-6), show6 = FALSE)
```

We can see that there are a few loci with P-values less than 1e-4, and
even some with P-values less than 1e-6.

In real datasets, it is wise to have a little wriggle-room: this
dataset may still contain duplicate samples, fish from another
population, or contaminated or degraded samples - and each of those
can contribute to the _loci_ having lower-than-expected P-values. So
for now, we will just remove loci with P < 1e-6. After the rest of the
cleaning stages, we can (and will!) check again for any loci that
might still be outlying.

```{r bluefinBadLoci, fig.show = 'hide'}
bluefin_1 <- bluefin[, pvals$pval4 > 1e-6]
pvals_1 <- check6and4(bluefin_1, thresh_pchisq_6and4 = c(1e-4, 1e-6), show6 = FALSE)
## pvals_1 is storage so we can check back on the same scores later
```

## finding and removing duplicate samples

Now, let's check for duplicate samples

```{r bluefinFindDups}
## look for duplicates in 'bluefin' and keep any pairs with less than 50
## different genotypes:
dups <- find_duplicates(bluefin_1, max_diff_loci = 50)
nrow(dups) ## 24 pairs of animals with fewer than 50 different genotypes
```

We've got 24 pairs of animals with fewer than 50 different genotypes -
they're almost certainly duplicates. When we have duplicate pairs, we
only need to remove one from each pair. `kinference` contains a
convenient function - `drop_dups_pairwise_equiv` - to help us do
exactly that:

```{r dropbluefindups}
bluefin_2 <- bluefin_1[ -c( drop_dups_pairwise_equiv( dups[,2:3])),]
```

## finding and removing samples from another population or species

Now, we need to find and remove any animals from another species or
population that might have made it into our dataset.

```{r bluefinIlglk}
ilglk <- ilglk_geno(bluefin_2)
abline(v = -1450) ## put our cutoff here
```

This pattern of `ilglk_geno` scores is pretty typical when most
samples are correctly identified, but there are a handful of animals
from a distantly-related stock or another species in the dataset:
there is a large, approximately Normally-distributed, spike at the
right-hand side of the plot, and a long tail strung out to the left,
containing the animals from another stock or species. In our case, any
animal with an ilglk score less than -1450 or so is not believable as
a member of the same stock as the main spike. Let's remove them.

```{r bluefinIlglk2}
bluefin_3 <- bluefin_2[ilglk > -1450,]
## now, does it look any better?
ilglk <- ilglk_geno(bluefin_3)
bluefin_3 <- kin_power(bluefin_3, k = 0.5) ## important but boring internal calculation - just do it.
```

## finding and removing cross-contaminated and degraded samples

Now, we need to find and remove cross-contaminated or degraded
samples.

```{r bluefinHetzes}
hetzrich <- hetzminoo_fancy(bluefin_3, "rich") ## a couple of outlying low samples here
hetzpoor <- hetzminoo_fancy(bluefin_3, "poor") ## also a couple here

bluefin_4 <- bluefin_3[hetzrich > 0.17,] ## remove the outlying hetzrich scores

hetzpoor <- hetzminoo_fancy(bluefin_4, "poor") ## Hey, the hetzpoor outliers are gone!
## (They were the same couple of fish as in hetzrich. No need to remove them now)
bluefin_5 <- prepare_PLOD_SPA(bluefin_4)
```

The tests for contaminated and degraded samples are both
likelihood-based, optimised for one or the other problem. However, as
in this case, there's often not much difference between these two test
statistics - fish that look abnormal on one also often look abnormal
on the other.

## how has removing bad fish impacted our apparent locus quality?

Remember how I mentioned that having bad _samples_ in the dataset can
make _loci_ look bad? Let's take a look at the P values for individual
loci we stored earlier, re-calculate them now that we've removed all
the bad fish, and see whether the loci look any better.

```{r bluefin6and4sMk2}
pvals_2 <- check6and4(bluefin_5, thresh_pchisq_6and4 = c(1e-4, 1e-6), show6 = FALSE)
lociPvals <- data.frame(pvals = c(pvals_2$pval4, pvals_1$pval4),
                        stage = c(rep("after", length(pvals_2$pval4)),
                                  rep("before", length(pvals_1$pval4)))
                        )
lm1 <- lm(pvals ~ stage, data = lociPvals)
summary(lm1)
```

Removing bad fish has caused a small, but highly
statistically-significant, improvement in the apparent quality of our
loci: the average locus in our dataset has seen its P-value increase
by 0.016. This is why we didn't want to be _too_ harsh in trimming
loci early in our cleaning pathway - we would have removed loci that
were actually fine, but looked bad because of the bad fish, and that
would have cost us power by reducing the number of good loci we use
for kin-finding. If you _really_ want to refine your locus-selection
to keep as many as possible, you might try a _very_ lenient cutoff for
locus quality, then look for very bad samples, then tighten the locus
quality cutoff, then look for bad samples again.

## kinfer the bluefin dataset

The `bluefin` dataset is now cleaned. We won't go through the full
kin-finding routine again here - it would be the same as for
`dropbears` (see **Section One**). I'll just show you the PLOD, so you
can see that kin-finding would work for `bluefin`, cleaned as it is.

```{r bluefinKinfer}
HSPs <- find_HSPs(bluefin_5, keep_thresh = -10)
PLOD_loghisto(HSPs)
```

This `PLOD_loghisto` shows a spike of unrelated pairs that is
well-characterised by its expected distribution, a spike in density
where we would expect to see half-siblings, and another spike in
density where we would expect full siblings or parent-offspring
pairs. From this, it should be clear that we can kinfer this dataset
using the same approach as in Section One, albeit with slightly less
power.

# Section Three: do we _really_ need to do all that cleaning?

```{r meanwhileOffscreen, include = FALSE}
## identify the animals or loci removed by each process; stick 'em
## in a data.frame; then subset bluefin by each.

## bad loci are in bluefin but not in bluefin_1
isBadLocus <- !bluefin@locinfo$Locus %in% bluefin_1@locinfo$Locus
badLoci <- bluefin@locinfo$Locus[isBadLocus]
## dups are in bluefin_1 but not bluefin_2
isDup <- !bluefin_1@info$Our_sample %in% bluefin_2@info$Our_sample
dups <- bluefin_1@info$Our_sample[isDup]
## bad ilglks are in bluefin_2 but not bluefin_3
isIlglk <- !bluefin_2@info$Our_sample %in% bluefin_3@info$Our_sample
ilglks <- bluefin_2@info$Our_sample[isIlglk]
## badhetz are in bluefin_3 but not in bluefin_4
isBadhetz <- !bluefin_3@info$Our_sample %in% bluefin_4@info$Our_sample
badhetz <- bluefin_3@info$Our_sample[isIlglk]

## now, do the 'all but one' removals
bluefin_exceptLoci <- bluefin[! bluefin@info$Our_sample %in% c(dups, ilglks, badhetz) ,]
bluefin_exceptLoci <- kin_power(bluefin_exceptLoci, k = 0.5)
bluefin_exceptLoci <- prepare_PLOD_SPA(bluefin_exceptLoci)


bluefin_exceptDups <- bluefin[! bluefin@info$Our_sample %in% c(ilglks, badhetz) ,
                              ! bluefin@locinfo$Locus %in% badLoci]
bluefin_exceptDups <- kin_power(bluefin_exceptDups, k = 0.5)
bluefin_exceptDups <- prepare_PLOD_SPA(bluefin_exceptDups)

bluefin_exceptIlglks <- bluefin[! bluefin@info$Our_sample %in% c(dups, badhetz) ,
                                ! bluefin@locinfo$Locus %in% badLoci]
bluefin_exceptIlglks <- kin_power(bluefin_exceptIlglks, k = 0.5)
bluefin_exceptIlglks <- prepare_PLOD_SPA(bluefin_exceptIlglks)

bluefin_exceptHetz <- bluefin[! bluefin@info$Our_sample %in% c(ilglks, dups) ,
                              ! bluefin@locinfo$Locus %in% badLoci]
bluefin_exceptHetz <- kin_power(bluefin_exceptHetz, k = 0.5)
bluefin_exceptHetz <- prepare_PLOD_SPA(bluefin_exceptHetz)
```

Offscreen, I've prepared versions of the `bluefin` dataset with
all-but-one of the cleaning steps. For example, `bluefin_exceptIlglk`
has had its loci checked, has had duplicates removed, had its
contaminated and degraded samples removed - but I've kept all the
samples that failed our check for animals from another population,
using `ilglk_geno`. I've also built `bluefin_exceptDuplicates`,
`bluefin_exceptHetzPoor`, `bluefin_exceptHetzRich`, and
`bluefin_except6and4`, each of which has skipped the obvious cleaning
step.  Here, we'll walk through the sorts of things you might see if
your data aren't properly cleaned.

## Duplicates

_In principle_, there are two reasons we don't want duplicate samples
in our kin-finding dataset: first, because our null distributions for
our various kin-finding statistics are calculated under the assumption
that each sample is drawn independently and at random from the sampled
population. Duplicate samples, especially if they account for a large
proportion of the dataset, will make those null distributions and
decisions based on them inaccurate. Second, if you follow the
kin-finding pathway shown in this vignette, and have duplicate samples
in your dataset, those dups will be mistaken for parent-offspring
pairs.

Duplicate pairs are very easy to spot: they look more closely related
than any true kin pair, so they appear far off to the right of the
`PLOD_loghisto` or `HSP_histo` plots. The `bluefin` dataset doesn't
have all that many duplicates, so the differences to the null
distributions aren't pronounced.

```{r keptDups}
dupHSPs <- find_HSPs(bluefin_exceptDups, keep_thresh = -10)
## PLOD_loghisto(dupHSPs) # use the PLOD_loghisto to pick the fullsib_cut
HSP_histo(dupHSPs, lb = 20, fullsib_cut = 150)
```

## Samples from another population or species

Samples from another population or species will look very unlike the
bulk of the other samples, so pairs with one animal from population A
(the intended population) and another from population B (the
unintended population) will tend to have very un-kin-like kinship
scores. On the other hand, pairs between two samples from population B
_may_ have very high kinship scores, especially if some loci have
alleles that are rare in population A but more common (or even fixed)
in population B.

```{r keptIlglks, fig.height = 11}
# par(mfrow = c(2,1))
ilglkHSPs <- find_HSPs(bluefin_exceptIlglks, keep_thresh = -10)
PLOD_loghisto(ilglkHSPs) # there's a spike out to the left of the plotting area
ilglkHSPs <- find_HSPs(bluefin_exceptIlglks, keep_thresh = -10, minbin = -250)
PLOD_loghisto(ilglkHSPs) # extend the plotting area (requires re-running find_HSPs)
```

Note both the bump out to the left of the unrelated pairs, and the
additional pairs scattered throughout the 'kin-like' regions of the
PLOD distribution, but not closely matching the expected peak locations
for any particular kin group except by accident.

## Contaminated or degraded samples

A few contaminated or degraded samples in a mostly-clean dataset
behave very similarly to samples from another population or species:
when paired with clean samples, they will look very dissimilar, but
when paired with each other, they can look very much like
kin.

The `bluefin` dataset did not have many cross-contaminated or degraded
samples, so we will demonstrate cross-contaminated and/or degraded
samples using a dataset from Andersen et al. (2019), available on
Dryad. We present the PLOD plots for this dataset with and without
cleaning for cross-contamination.


```{r tokelauOffscreen, include = FALSE}
if( FALSE){ # FOR NOW since the file does not exist..!
load("~/R/localPackages/kinference/kinference/data/yellowfin.rda") ## loads 'tokelau_snpgeno'
load("~/Dropbox/CSIRO/contamStat/yellowfin_probablyContam.rda") ## use 'ttlc' as a sample ID
tokelau_snpgeno@info$probablyContam <- info$probablyContam[match(tokelau_snpgeno@info$ttlc, info$ttlc)]
tokelau_snpgeno@info$ttlcCheck <- info$ttlc[match(tokelau_snpgeno@info$ttlc, info$ttlc)]

tokelau_clean <- tokelau_snpgeno[tokelau_snpgeno@info$probablyContam == FALSE,]
tokelau_all <- tokelau_snpgeno

## tokelau_all@locinfo$pbonzer <- re_est_ALF(tokelau_all)@locinfo$pambig

## checking for badly-behaved loci
pvals <- check6and4(tokelau_all, thresh_pchisq_6and4 = c(1e-4, 1e-6), show6 = FALSE)
## allelic dropout is a big problem with the ToK data: see
## over-representation of OO, AAO and BBO, and under-representation of
## AB

tokelau_all_1 <- tokelau_all[, pvals$pval4 > 0.2]

## finding and removing duplicate samples

## look for duplicates in 'tokelau_all' and keep any pairs with less than 50
## different genotypes:
dups <- find_duplicates(tokelau_all_1, max_diff_loci = 500, show_plot = FALSE)
nrow(dups) ## no duplicates
tokelau_all_2 <- tokelau_all_1 ## nothing to remove

## finding and removing samples from another population or species
ilglk <- ilglk_geno(tokelau_all_2) ## doesn't align to priors _at all_. Maybe re-do later, for didactic reasons?

tokelau_all_3 <- tokelau_all_2
tokelau_all_3 <- kin_power(tokelau_all_3, k = 0.5)

## finding and removing cross-contaminated and degraded samples

hetzrich <- hetzminoo_fancy(tokelau_all_3, "rich")
hetzpoor <- hetzminoo_fancy(tokelau_all_3, "poor")

tokelau_all_4 <- tokelau_all_3

tokelau_all_5 <- prepare_PLOD_SPA(tokelau_all_4)

contamHSPs <- find_HSPs(tokelau_all_5, keep_thresh = -10)
PLOD_loghisto(contamHSPs) # wait, that looks like _cousins_ in the tokelau data
## ok, the tokelau data aren't a great example to use - too much allelic dropout; not enough contamination
} # if FALSE
```








```{r keptHetz, fig.height = 11}
par(mfrow = c(2,1))
hetzHSPs <- find_HSPs(bluefin_5, keep_thresh = -10)
PLOD_loghisto(hetzHSPs, main = "bluefin") # there's a spike out to the left of the plotting area
hetzHSPs <- find_HSPs(bluefin_exceptHetz, keep_thresh = -10)
PLOD_loghisto(hetzHSPs, main = "bluefin_exceptHetz") # extend the plotting area (requires re-running find_HSPs)
```
